On The Difficulty of Computing Non-Commutative Signature Gröbner Bases

### 2022-02-02 article_issac.tex
Modifications générateurs
Quelques corrections sur la partie Langage régulier.

  Now, let us show that leading monomials of syzygies of the form $\varepsilon_{11}a^nb^me$ are such that $n = m$.
  Let $\sigma$ be such a syzygy and let $G$ be the graph whose vertices are elements of $\left\langle X \right\rangle$, and there is an edge between vertices of the form $lXr$ and $lYr$ as soon as $lg_ir$ occurs with a nonzero coefficient in $\sigma$ and $g_i=X - Y$. Let $W$ be the set of vertices of the following form:
  \begin{align*}
    & ssa^nb^ne, \\
    & tc^pa^qxa^rd^nb^me, \\
    & tc^pa^qd^rxd^nb^me, \\
    & tc^pa^qya^rd^nb^me, \\
    & tc^pa^qd^ryd^nb^me, \\
    & tc^pza^qb^re, \\
    & tc^pd^qzb^re
  \end{align*}
  Each vertex in $W$ is related to at most two vertices, which also belong to $W$. When an undirected graph is such that each vertex is related to at most two vertices, each connected component is either a chain or cycle, so that the connected component of $ssa^nb^me$ in $G$ is either a chain or a cycle.
  But since $\sigma$ is a syzygy, this connected component has to be a cycle, which implies $n = m$.

  Now, we can conclude that the set $X$ of leading monomials of syzygies of $F$ does not form a regular language.
  Indeed, by way of contradiction, assume that $X$ is a regular language, so that its
  intersection with the regular language $\{g_{11}a^nb^me \;|\; n, m \geq 1\}$ also forms regular language.
  We reach a contradiction since we proved that this intersection is $\{g_{11}a^nb^ne \;|\; n \geq 1 \}$, which does not form a regular language.
\end{proof}

Notice that $F=\{g_1,\dots,g_{11}\}$ has no reason to have a finite Gröbner basis. However, we construct a new set $F'$ by adding to $F$ the polynomial $\infty$ and all polynomials of the form $\infty-\alpha$, for letters $\alpha$ different from $\infty$. Then $F'$ admits the reduced Gröbner basis which contains each letter.
Now, let $X$ and $Y$ be the sets of leading monomials of syzygies of $F$ and $F'$, respectively, and let 
$$Z=\{l\varepsilon_ir \in \MM(\mathscr{F}_r) \;|\; l \LM(g_i) r \mbox{ does not contain } \infty\}.$$
Since $X=Y\cap Z$ \fixmecc{justifier cette formule}, $Z$ is a regular language and $X$ is not a regular language, $Y$ is not a regular language. Hence, we have constructed a set $F'$ of polynomials that admit a finite Gröbner basis whose set of leading monomials of syzygies does not form a regular language.

@@

>>>>

  Now, let us show that leading monomials of syzygies of the form $\varepsilon_{11}a^nb^me$ are such that $n = m$.
  Let $\sigma$ be such a syzygy and let $G$ be the graph whose vertices are elements of $\left\langle X \right\rangle$, and there is an edge between vertices of the form $l\omega r$ and $l\omega'r$ as soon as $lg_ir$ occurs with a nonzero coefficient in $\sigma$ and $g_i=\omega - \omega'$. Let $W$ be the set of vertices of the following form:
  \begin{align*}
    & ssa^nb^ne, \\
    & tc^pa^qxa^rd^nb^me, \\
    & tc^pa^qd^rxd^nb^me, \\
    & tc^pa^qya^rd^nb^me, \\
    & tc^pa^qd^ryd^nb^me, \\
    & tc^pza^qb^re, \\
    & tc^pd^qzb^re
  \end{align*}
  Each vertex in $W$ is related to at most two vertices, which also belong to $W$. When an undirected graph is such that each vertex is related to at most two vertices, each connected component is either a chain or cycle, so that the connected component of $ssa^nb^me$ in $G$ is either a chain or a cycle.
  But since $\sigma$ is a syzygy, this connected component has to be a cycle, which implies $n = m$.
  Let $\mathscr{C}$ be defined by:
$$\mathscr{C}=\{l\varepsilon_ir \in \MM(\mathscr{F}_r) \;|\; l \LM(g_i) r \mbox{ does not contain } \infty\}.$$
   Let $\mathscr{A}$ be the set of leading monomials of syzygies of $F$.
   Now, we can conclude that the set $\mathscr{A} \cap \mathscr{C}$ does not form a regular language.
  Indeed, by way of contradiction, assume that $\mathscr{A}\cap \mathscr{C}$ is a regular language. Then its
  intersection with the regular language $\{g_{11}a^nb^me \;|\; n, m \geq 1\}$ also forms regular language.
  We reach a contradiction since we proved that this intersection is $\{g_{11}a^nb^ne \;|\; n \geq 1 \}$, which does not form a regular language.
\end{proof}

Notice that $F=\{g_1,\dots,g_{11}\}$ has no reason to have a finite Gröbner basis. However, we construct a new set $F'$ by adding to $F$ the polynomial $\infty$ and all polynomials of the form $\infty-\alpha$, for letters $\alpha$ different from $\infty$. Then $F'$ admits the reduced Gröbner basis which contains each letter.
Now, let $\mathscr{B}$ be the sets of leading monomials of syzygies of $F'$.
As our ideal is homogeneous,
the set of the
leading monomials of the syzygies
is the same as the set of the leading
monomials of the homogeneous syzygies.
Thanks to the way we defined our order,
if $\sigma \in Syz(F')$ is homogeneous,
and
if $\infty$ appears in a $l \LM(g_i) r$,
for some $l \varepsilon_i r$ 
a term of $\sigma.$ 
then 
$\infty$ appears in $l' \LM(g_j) r'$
where $l' \varepsilon_j r'$ is the $\LM$
of $\sigma.$
Thus $\mathscr{A} \cap \mathscr{C}=\mathscr{B} \cap \mathscr{C}$. Since $\mathscr{C}$ is a regular language and $\mathscr{A} \cap \mathscr{C}$ is not a regular language, $\mathscr{B}$ is not a regular language. Hence, we have constructed a set $F'$ of polynomials that admit a finite Gröbner basis whose set of leading monomials of syzygies does not form a regular language.

@@

\begin{align*}
g_1 & := ssa - t\alpha x \\ 
g_2 & := xa - ax \\ 
g_{11} & := x\beta-\beta x\\ 
g_{12} & := xb-\beta y \\ 
%g_3 & := xbb - \beta yb\\ 
% Inutile, redondante avec g_12
g_4 & := yb-by\\ 
g_{13} & := y\gamma-\gamma y \\
g_5 & := ycc-\gamma zc \\ 
g_6 & :=  \gamma z-z\gamma\\ 
g_7 & := bz-zb \\
g_8 & := \beta z-z\beta \\
g_9 & := az-za \\ 
g_{10} & := \alpha za-\alpha\alpha x  \\ 
g_{19} & := yce - \gamma oe \\
g_{14} & := \gamma o-oc \\
%g_{15} & := oc-\gamma o \\ 
% Inutile, redondante avec g_14
g_{16} & := \beta o-ob \\ 
g_{17} & := \alpha o-oa \\ 
g_{18} & := to-ss
\end{align*}
